---
title: 'M Carlin: Third Deliverable'
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}
setwd("C:/Users/CarlinML/DACSS-690R/Third_Deliverable")
getwd()
```

## CLERKSHIP GRADES DATASET - MERGING

# Cleaned/formatted clerkship grade data

```{r}
ClerkshipGrades <- readRDS(gzcon(url("https://github.com/DACSS-690R/Second_Deliverable/raw/main/DataCleanAndFormatted/ClerkshipGrades_formatted.RDS")))
names(ClerkshipGrades)[names(ClerkshipGrades) == 'Session...6'] <- 'Session'
names(ClerkshipGrades)[names(ClerkshipGrades) == 'Session...7'] <- 'Session_Code'
names(ClerkshipGrades)[names(ClerkshipGrades) == 'Location'] <- 'Location_Code'
ClerkshipGrades$Location_new = ClerkshipGrades$Location_Description
str(ClerkshipGrades) # 6427 rows
```

# Gender and ethnicity info

```{r}
library(readxl)
folder="Data"
fileName="Gender_Ethnicity_info.xlsx"
fileToRead=file.path(folder,fileName)
fileToRead
Gender_Ethnicity=read_xlsx(fileToRead)
str(Gender_Ethnicity) # 1004 rows
```

# merge ClerkshipGrades with Gender_Ethnicity using left join

```{r}
ClerkshipGrades2 <- merge(ClerkshipGrades,Gender_Ethnicity,all.x=TRUE, by.x="ID", by.y="ID")
str(ClerkshipGrades2) # 6427 rows 
```

# Location info

```{r}
folder="Data"
fileName="Location_info.xlsx"
fileToRead=file.path(folder,fileName)
fileToRead
Locations=read_xlsx(fileToRead)
str(Locations) # 22 rows
```

# merge ClerkshipGrades_with_GenderEthnicity with Locations using fuzzy merge

# basic left merge

```{r}
str(merge(ClerkshipGrades2,Locations, by.x='Location_new', by.y='Location'))
# 1877 matched rows
```

# locations only in ClerkshipGrades_GenderEthnicity file

```{r}
onlyInGrades=sort(setdiff(ClerkshipGrades2$Location_new,Locations$Location))
onlyInGrades
```

# locations only in Locations file

```{r}
onlyInLocations=sort(setdiff(Locations$Location, ClerkshipGrades2$Location_new))
onlyInLocations
```

```{r}
library(stringdist)
stringdist::stringdist(onlyInGrades[1],onlyInLocations,method = 'jaccard')
```


```{r}
TheMatch<-character()
mindist<-integer()
sortedmatches<-character()

for (i in 1:length(onlyInGrades) ) {
  allDistances<-stringdist::stringdist(onlyInGrades[i],onlyInLocations,method = 'jw')
  mindist[i]=min(allDistances)
  TheMatch[i]<-onlyInLocations[which.min(allDistances)]
}
fuzzyOutput=data.frame(input=onlyInGrades,possible_match=TheMatch,divergence=mindist,  stringsAsFactors = F)

fuzzyOutput[order(-fuzzyOutput$divergence),]
```

# keep all matches from above even though they aren't all completely correct; they will be corrected in subsequent steps

```{r}
fuzzyMatch_notperfect=fuzzyOutput[order(-fuzzyOutput$divergence),]
row.names(fuzzyMatch_notperfect)=NULL
fuzzyMatch_notperfect
```

# Manually change the 3 incorrect matches

```{r}
fuzzyMatch_notperfect[c(1,2,11),'input']
changes=c('Baystate Medical Center',"UMMHC Hospitals","Berkshire Medical Center")
fuzzyMatch_notperfect[c(1,2,11),'possible_match']=changes

fuzzyMatch_notperfect
fuzzyMatch_Perfect_part1=fuzzyMatch_notperfect[,-3] # without 3rd column
```


```{r}
##fuzzyMatch_Perfect_part2?
forPart2=intersect(Locations$Location, ClerkshipGrades2$Location_new)
fuzzyMatch_Perfect_part2=data.frame(cbind(input=forPart2,possible_match=forPart2))
fuzzyMatch_Perfect_part2
```

```{r}
# concat
fuzzyMatch_Perfect=rbind(fuzzyMatch_Perfect_part1,fuzzyMatch_Perfect_part2)
names(fuzzyMatch_Perfect)=c('Location_new','location_main')
```

```{r}
ClerkshipGrades_withLoc=merge(ClerkshipGrades2,fuzzyMatch_Perfect, by='Location_new',all.x = T)
str(ClerkshipGrades_withLoc)
```


# Pivot wide to long

```{r}
ClerkshipGrades_wide <- ClerkshipGrades_withLoc %>% select(-c(Location_new, X, Title, Session, Location_Code:Final_Letter, SPE_Letter, NBME_Letter, OSCE_Letter, gender, ethnicity))
names(ClerkshipGrades_wide) = gsub(pattern = "_Letter2", replacement = "_Letter", x = names(ClerkshipGrades_wide))
ClerkshipGrades_wide <- ClerkshipGrades_wide %>% relocate(location_main, .before = Final_Numeric)
ClerkshipGrades_wide

library(tidyr)
ClerkshipGrades_long <- ClerkshipGrades_wide %>% 
  gather(v, value, Final_Numeric:OSCE_Letter) %>% 
  separate(v, c("var", "col")) %>% 
  arrange(ID) %>% 
  spread(col, value)
ClerkshipGrades_long
str(ClerkshipGrades_long) # 25708 rows

# remove rows containing one or more NAs
ClerkshipGrades_long <- na.omit(ClerkshipGrades_long) # 20181 rows
```

